---
title: "Reproducible Research: Peer Assessment 1, Keith Maclaren"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r echo=TRUE}
myact <- read.csv("activity.csv")
str(myact)
```
Convert interval to a factor for analysis
```{r echo=TRUE}
myact$interval <- factor(myact$interval)
str(myact)
```
## What is mean total number of steps taken per day?
Calculate Total steps taken per day:
```{r echo=TRUE}
newdf <- aggregate(list(total_steps = myact$steps), list(date = myact$date), sum, na.rm=TRUE)
```
Histogram Plot of Total steps taken per day:
```{r echo=TRUE}
hist(newdf$total_steps)
abline(v=mean(newdf$total_steps), lty=3, col="blue")                   # draw a blue line thru the mean  
abline(v=median(newdf$total_steps), lty=4, col="red")                  # draw a red line thru the median  
text(mean(newdf$total_steps),25,labels="mean", pos=4, col="blue")       
text(mean(newdf$total_steps),23,labels="median", pos=4, col="red")  
```

Mean & Median steps taken per day:
```{r echo=TRUE}
summary(newdf$total_steps)
```


## What is the average daily activity pattern?
```{r echo=TRUE, fig.width=10}
newdf2 <- aggregate(list(avg_steps = myact$steps), list(interval=myact$interval), mean, na.rm=TRUE)
plot.new()
plot(newdf2$interval, newdf2$avg_steps, type="l")
lines(newdf2$interval, newdf2$avg_steps, col="black")
maxInt <- newdf2[which.max(newdf2$avg_steps),1]
abline(v=maxInt, lty=4, col="red")
```

## Imputing missing values
Calculate and report the total number of missing values

Strategy for filling in missing values = mean for that interval across all days

Create a new dataset with missing values filled in

Create a histogram of new dataset

Show mean/median of new dataset

Comparison from previous dataset

Impact of imputing values = ____________________


## Are there differences in activity patterns between weekdays and weekends?

New factor variable for weekday & weekend (use weekday() function against date variable coerced to datetime class)

Panel plot comparing activity plot by weekday & weekend