---
title: "Reproducible Research: Peer Assessment 1, Keith Maclaren"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r echo=TRUE}
myact <- read.csv("activity.csv")
str(myact)
```
Convert interval to a factor for analysis
```{r echo=TRUE}
myact$interval <- factor(myact$interval)
str(myact)
```
## What is mean total number of steps taken per day?
Calculate Total steps taken per day:
```{r echo=TRUE}
newdf <- aggregate(list(total_steps = myact$steps), list(date = myact$date), sum, na.rm=TRUE)
```
Histogram Plot of Total steps taken per day:
```{r echo=TRUE}
hist(newdf$total_steps)
abline(v=mean(newdf$total_steps), lty=3, col="blue")                   # draw a blue line thru the mean  
abline(v=median(newdf$total_steps), lty=4, col="red")                  # draw a red line thru the median  
text(mean(newdf$total_steps),25,labels="mean", pos=4, col="blue")       
text(mean(newdf$total_steps),23,labels="median", pos=4, col="red")  
```

Mean & Median steps taken per day:
```{r echo=TRUE}
summary(newdf$total_steps)
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
