---
title: "Reproducible Research: Peer Assessment 1, Keith Maclaren"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data - csv file contained in zip file
```{r echo=TRUE}
myact <- read.csv("activity.csv")
str(myact)
```
Convert interval to a factor for analysis
```{r echo=TRUE}
myact$interval <- factor(myact$interval)
str(myact)
```
## What is mean total number of steps taken per day?
Calculate Total steps taken per day:
```{r echo=TRUE}
newdf <- aggregate(list(total_steps = myact$steps), list(date = myact$date), sum, na.rm=TRUE)
```
Histogram Plot of Total steps taken per day:
```{r echo=TRUE}
hist(newdf$total_steps)
abline(v=mean(newdf$total_steps), lty=3, col="blue")                   # draw a blue line thru the mean  
abline(v=median(newdf$total_steps), lty=4, col="red")                  # draw a red line thru the median  
text(mean(newdf$total_steps),25,labels="mean", pos=4, col="blue")       
text(mean(newdf$total_steps),23,labels="median", pos=4, col="red")  
```

Mean & Median steps taken per day:
```{r echo=TRUE}
summary(newdf$total_steps)
```


## What is the average daily activity pattern?
```{r echo=TRUE, fig.width=10}
newdf2 <- aggregate(list(avg_steps = myact$steps), list(interval=myact$interval), mean, na.rm=TRUE)
plot.new()
plot(newdf2$interval, newdf2$avg_steps, type="l")
lines(newdf2$interval, newdf2$avg_steps, col="black")
maxInt <- newdf2[which.max(newdf2$avg_steps),1]
abline(v=maxInt, lty=4, col="red")
```

## Imputing missing values
Calculate and report the total number of missing values
```{r echo=TRUE}
cat("total records: ", nrow(myact))
cat("total NAs: ",sum(is.na(myact)))
```
Strategy for filling in missing values = mean for that interval across all days

Create a new dataset with missing values filled in
```{r echo=TRUE}
newact <- myact
# convert date to datetime class
newact$date <- as.Date(newact$date)
# loop through newact
for (i in 1:nrow(newact)) {
#   if find NA in newact$steps, replace with newdf2$steps (average) for matching interval
    if (is.na(newact[i,1])) {
        newact[i,1] <- newdf2[which(newdf2$interval == newact[i,3]), 2]
    }
}

```

Create a histogram of new dataset
```{r echo=TRUE}
newdf3 <- aggregate(list(total_steps = newact$steps), list(date = newact$date), sum, na.rm=TRUE)
hist(newdf3$total_steps)
abline(v=mean(newdf3$total_steps), lty=3, col="blue")                   # draw a blue line thru the mean  
abline(v=median(newdf3$total_steps), lty=4, col="red")                  # draw a red line thru the median  
text(mean(newdf3$total_steps),25,labels="mean", pos=4, col="blue")       
text(mean(newdf3$total_steps),23,labels="median", pos=4, col="red")  
```

Show mean/median of new dataset
```{r echo=TRUE}
summary(newdf3$total_steps)
```
Comparison from previous dataset
 - strengthens the mean, to the point the median == mean


## Are there differences in activity patterns between weekdays and weekends?

New factor variable for weekday & weekend (use weekdays() function against date variable)
```{r echo=TRUE}
newact$daytype <- ifelse(weekdays(newact$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")
head(newact, 2000)
```
Panel plot comparing activity plot by weekday & weekend